---
title: "Data Science HW10, Max Greene"
output: html_notebook
---
#Question #1



#Question #2

```{r}
library(ISLR)
```


```{r}
head(Weekly)
```

##(a)
Plots of "Today" and "Volume" with respect to time, each entry assumed to be a day, with adjacent rows being adjacent days.
```{r}
ggplot(Weekly,mapping = aes(x = 1:length(Weekly$Year),y=Today)) +   geom_point()

ggplot(Weekly,mapping = aes(x=1:length(Weekly$Volume),y=Volume))+
  geom_point()
```

It is apperant in this dataset that "Today" entries are fairy constant, while "Volume" entries gradually increase, followed by a sharp increase and a sharp decrease.
```{r}
summary(Weekly %>% dplyr::select(-Lag1,-Lag2,-Lag3,-Lag4,-Lag5))
```

##(b)
```{r}
weekly_df <- mutate(Weekly, dir_bin = as.integer(Direction)-1)
head(weekly_df)

ggplot(weekly_df,mapping = aes(x=1:length(weekly_df$dir_bin),y=dir_bin)) + geom_point() + stat_smooth(method = "lm")
```

```{r}
weekly_glm <- glm(
  dir_bin~Today:Lag1+Today,
  family=binomial,data=weekly_df)
summary(weekly_glm)
```

After checking many combinations of variables, Today in combination with Today:Lag1 is the best predictor of tomorrow's outcome mobing up or down.

##(c)
```{r}
prediction <- predict(weekly_glm,type="response")
prediction <- ifelse(prediction<=0.1,0,1)
table(weekly_df$dir_bin,prediction)
```

This confusion matrix says the following things about the glm:
1) 483 decreases were correctly predicted
2) 1 decrease was incorrectly predicted as an increase
3) 603 increases were correctly predicted
4) 2 increases were incorrectly predicted as a decrease

##(d)
Create test and train data sets, where training sets are up until 2008 and test sets are all of 2009 and 2010.
```{r}
weekly_df_train <- mutate(Weekly, dir_bin = as.integer(Direction)-1)
weekly_df_test <- dplyr::filter(weekly_df_train, Year > 2008)
weekly_df_train <- dplyr::filter(weekly_df_train, Year < 2009)
```

```{r}
weekly_glm <- glm(dir_bin~Lag2,family=binomial,data=weekly_df_train)
summary(weekly_glm)
```

```{r}
prediction <- predict(weekly_glm,weekly_df_test,type="response")
prediction <- ifelse(prediction<=0.5,0,1)
temp <- table(weekly_df_test$dir_bin,prediction)
temp
```

Fraction of percentage correct:
```{r}
(sum(diag(temp)))/(sum(temp))*100
```

##(e)
Create/reset test and train data sets, where training sets are up until 2008 and test sets are all of 2009 and 2010.
```{r}
weekly_df_train <- mutate(Weekly, dir_bin = as.integer(Direction)-1)
weekly_df_test <- dplyr::filter(weekly_df_train, Year > 2008)
weekly_df_train <- dplyr::filter(weekly_df_train, Year < 2009)
```

```{r}
weekly_lda <- lda(dir_bin~Lag2,family=binomial,data=weekly_df_train)
weekly_lda
```

```{r}
prediction <- predict(weekly_lda,weekly_df_test,type="response")
table(weekly_df_test$dir_bin,prediction$class)
```


##(f)
Create/reset test and train data sets, where training sets are up until 2008 and test sets are all of 2009 and 2010.
```{r}
weekly_df_train <- mutate(Weekly, dir_bin = as.integer(Direction)-1)
weekly_df_test <- dplyr::filter(weekly_df_train, Year > 2008)
weekly_df_train <- dplyr::filter(weekly_df_train, Year < 2009)
weekly_df_train <- head(weekly_df_train,length(weekly_df_test$Today))
```

```{r}
#weekly_df_train$dir_bin
temp1 <- as.matrix(dplyr::select(weekly_df_train,Lag2))
temp2 <- as.matrix(dplyr::select(weekly_df_train,dir_bin))
weekly_knn <- knn(temp1,temp1,temp2,k=1)
table(temp2,weekly_knn)
```

##(g)
Create test and train data sets, where training sets are up until 2008 and test sets are all of 2009 and 2010.
```{r}
weekly_df <- mutate(Weekly, dir_bin = as.integer(Direction)-1)
```

```{r}
weekly_glm <- glm(dir_bin ~ Today + Today:Lag1
                  ,family=binomial,data=weekly_df)
summary(weekly_glm)
```

```{r}
prediction <- predict(weekly_glm,weekly_df_test,type="response")
prediction <- ifelse(prediction<=0.5,0,1)
temp <- table(weekly_df_test$dir_bin,prediction)
temp
```

Fraction of percentage correct:
```{r}
(sum(diag(temp)))/(sum(temp))*100
```

This was the most accurate formula for the `glm` that I could find, which makes sense because it includes today's value and a combination of the previous day and today, likely capturing short-term trends.






