---
title: "R Notebook on BaseBall Salaries"
output: html_notebook
---

# Introduction
Dis nitebook for dem BaseBall Salaries

## Introduction Subsection
Dis subsection for dem section subsections

```{r}
library(tidyverse);
library(reshape2);
```

Now I will load the data.
```{r}
salary_data <- read.csv("Data/Salaries.csv")
```

Get the size fo the dataframe.
```{r}
dim(salary_data)
```

Here are the first few rows.
```{r}
head(salary_data)
```

Here are the last few rows.
```{r}
tail(salary_data)
```

Obtain the names of the columns.
```{r}
names(salary_data)
```

Obtain summary of stats.
```{r}
summary(salary_data)
```

Make a bar plot indicating the various teams.
```{r}
salary_data %>% ggplot() + geom_bar(mapping = aes(x = teamID))+ coord_flip()
```


# Examine Salary by Year
Consider average player salary by year
```{r}
avgSalByYear <- summarise(group_by(salary_data,yearID),mean_sal = mean(salary))
avgSalByYear
ggplot(avgSalByYear) + geom_line(mapping = aes(x = yearID,y = mean_sal))
```





















