---
title: "Historical Ethereum Price Exploratory Data Analysis"
output: html_notebook
---

Load necessary libraries:
```{r}
library(tidyverse);
#may need to install:
library(crypto);
library(zoo);
```

#Import Data

Import data with `crypto` package using my function `GetCoinHistory()` from the `gatherData.R` file:
```{r}
data <- GetCoinHistory("ETH")
```

#Organize Data

Add 10 and 50 day moving average of open prices with my function `mutateRoll()` from the `mutateData.R` file:
```{r}
dataMA <- mutateRoll(data,data$open,c(10,50))
```

Add Yesterday's open and close price to data with my function `mutateYesterday()` from the `mutateData.R` file:
```{r}
dataMA <- mutateLast(dataMA,data$open,"openLast")
dataMA <- mutateLast(dataMA,data$close,"closeLast")
```

Add Tomorrow's open price to data with my function `mutateLast()` from the `mutateData.R` file:
```{r}
dataMA <- mutateNext(dataMA,data$open,"openNext")
dataMA <- mutateNext(dataMA,data$close,"closeNext")
```
These values, the open and close price for tomorrow's market, will serve as the training values for our model. 

#Plot Data
I have written four plotting functions in my `plotData.R` file:
`plotOpen()`, `plotClose()`, `plotHigh()`, `plotLow()`

```{r}
plotOpen(dataMA)
plotClose(dataMA)
plotHigh(dataMA)
plotLow(dataMA)
```

#Moving Forward w/ Model
Now that the data has moving averages as well as past and previous values for open and close, it can be implemented in the form of a model. 
Yesterday's open, yesterday's close, today's open, today's close, historically the open and close of the next day, 10-day moving averages and 50-day moving averages can now be predictors used in the model, which could have a variety of useful outputs in trading. Namely, the difference between tomorrow's open and close values would be a useful prediction for traders, because it would indicate a predicted large shift in the market.



