---
title: "StatsIntro"
author: "Maxwell Greene"
date: "October 16, 2019"
output: pdf_document
---

#Importing data to table from clipboard:
First, copy the entire dataset to the clipboard, including the header row.
Then run the following code to assign it to a variable:
(I've commented this portion out since I will be loading from a file when running it in this notebook.)
```{r}
#dataset <- read.table("clipboard",header=TRUE)
```

I've saved the dataset in a file in the same folder as this notebook, so I can use the read.table command from the file instead of the clipboard.
```{r}
dataset <- read.table("dataset1.csv",
                      header=TRUE,sep=",")
```

###A note about notation (the "<-" operator, specific to R):
You can think of the "<-" operator as "=". More accurately, though, we should say the LHS "gets" or "is assigned" the value of the RHS. There are subtle differences between "<-" and "=" (I'll show one below), but "<-" is more flexible and should be used by default.
```{r}
sum(x=1,2); exists('x')
sum((x=1),2); exists('x')
sum(y<-1,2); exists('y')
rm(x,y)
#Notice the equals assignment is only stored when in perenthesis,
#whereas the y assignment need not be in perenthesis.
```

#Creating Plots of Difference (Start - End)

Create a bixplot for diff vs. vision:
```{r,fig.width=6,fig.height=3}
boxplot(diff~Vis,data=dataset,
        main="Diff vs. Vision",
        xlab="Can this person see?",
        ylab="Difference (start-end)")
```

Create a scatterplot for diff vs. height:
```{r,fig.width=6,fig.height=3}
plot(x=dataset$Height,y=dataset$diff,
     main="Height vs. Diff",
     xlab="Height (cm)",
     ylab="Difference (start-end)")
```

Create a plot of diff vs. person, an easy way to visualize the distribution of differences among individuals:
```{r,fig.width=6,fig.height=3}
plot(x=dataset$Person,y=dataset$diff)
```

#Create Plots of Walking (Tstart-End)
Create a bixplot for walking vs. vision:
```{r,fig.width=6,fig.height=3}
boxplot(Walking~Vis,data=dataset,
        main="Walking vs. Vision",
        xlab="Can this person see?",
        ylab="Walking (Tstart-end)")
```

Create a scatterplot for walking vs. height:
```{r,fig.width=6,fig.height=3}
plot(x=dataset$Height,y=dataset$Walking,
     main="Height vs. Walking",
     xlab="Height (cm)",
     ylab="Walking (Tstart-end)")
```






