---
title: "EDA w/ Focus on Ratios"
output:
  html_notebook: default
  html_document:
    df_print: paged
---

This notebook is meant for importing, tidying and formatting my data to be modelled. Here I will focus on ratios, effectively ignoring the absolute value of any open or close price for each day.

Load necessary libraries:
```{r}
library(tidyverse);
library(crypto);
library(zoo);
```

#Import Data

Import data with `crypto` package using my function `GetCoinHistory()` from the `gatherData.R` file:
```{r}
data <- GetCoinHistory("ETH")
```

#Organize Data Using my .R Functions
Add 10 and 50 day moving average of open prices with my function `mutateRoll()` from the `mutateData.R` file:
```{r}
data <- mutateRoll(data,data$open,c(5,10,25))
```

Add Yesterday's open and close price to data with my function `mutateLast()` from the `mutateData.R` file:
```{r}
data <- mutateLast(data,data$open,"openLast")
data <- mutateLast(data,data$close,"closeLast")
data <- mutateLast(data,data$high,"highLast")
data <- mutateLast(data,data$low,"lowLast")
data <- mutateLast(data,data$volume,"volumeNext")
```

Add Tomorrow's open price to data with my function `mutateNext()` from the `mutateData.R` file:
```{r}
data <- mutateNext(data,data$open,"openNext")
data <- mutateNext(data,data$close,"closeNext")
data <- mutateNext(data,data$high,"highNext")
data <- mutateNext(data,data$low,"lowNext")
data <- mutateNext(data,data$volume,"volumeNext")
```

Now, convert all applicable columns to ratios with my function `convertRatio()` from the `mutateData.R` file:
```{r}
columnNames <- c("close","low","high","openLast","closeLast","highLast","lowLast","openNext","closeNext","highNext","lowNext","ma1","ma2","ma3")
dataRatio <- convertRatio(data,data$open,columnNames)
```


#Model Evaluation
##Statistical Analysis of Model
**Incomplete, will continue model analysis**
```{r}
library(boot)
formula <- as.formula("openNext~open+close+ma1")
glm.fit <- glm(formula=formula,data=dataRatio)
glm.fit
#cv.glm(dataRatio,glm.fit,K=10)$delta
```

##Visualizing Accuracy
The following graph is a visual representation of the accuracy of the model, showing the percentage of predictions within the given percentage of the correct test data value. This was done using my function `glmVect()` in the file `modelData.R`. Note: The function `glmPlot()` in the same file performs the same computations, but saves a .png file of the plot, rather than returning a vector.
```{r}
formula <- as.formula("openNext~open+close+volume+high+low+market+ma1+ma2+ma3")
percents <- seq(.01,2,.01)
result <- as.vector(glmVect(1000,200,percents,formula,forNum=100,data1=dataRatio))
ggplot(data=data.frame(result=result,value=percents)) + geom_point(mapping = aes(x=value,y=result))
```










